---
title: "algalblooms_analysis"
author: "Kate Becker"
date: "2023-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(here)
library(janitor)
library(dplyr)
library(lubridate)
library(sf)
library(terra)
library(spData)
library(spDataLarge)
library(maps)
library(stringr)
library(viridis)
library(gstat)
library(sp)
library(automap)
library(tufte)
library(feasts)
library(forecast)
library(tsibble)
library(tsibble)
library(feasts)
library(tmap)
library(zoo)
library(spData)

```

```{r}
rm(list = ls())
here::i_am("algae_temp_analysis.Rmd")
setwd(here())
```

```{r}
algal_events <- read_csv("/Users/katebecker/Documents/Bren/Fall_Q/EDS_222/ final_proj/algalblooms_sb/data/edi/event.csv")
algal_measure <- read_csv("/Users/katebecker/Documents/Bren/Fall_Q/EDS_222/ final_proj/algalblooms_sb/data/edi/extendedmeasurementorfact.csv")
algal_occurence <- read_csv("/Users/katebecker/Documents/Bren/Fall_Q/EDS_222/ final_proj/algalblooms_sb/data/edi/occurrence.csv")
santa_barbara <- st_read(here("data", "tl_2019_06083_faces", "tl_2019_06083_faces.shp"))
sst <- readr::read_csv(here("data", "sst", "sst_original_data.csv"))
```

## Linear Regression for Sea Surface Temp and Date (15 Years)

```{r}
sst_filter <- sst %>%
  mutate(year = format(date, "%Y"))

updated_sst <- sst_filter %>%
  filter(year >= 2007 & year <= 2022)
```

```{r}
linear_model <- lm(temp ~ date, data = updated_sst)
slope <- coef(linear_model)[2]

#we've all see this general trend, showing seasonal oscillations
ggplot(data = updated_sst, aes(x = date, y = temp)) +
  geom_point(color = "#726161", size = 0.25) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Time", 
       y = "Sea Surface Temperature (°C)") + 
  ggtitle("Sea Surface Temperature Trend Over the Last 15 Years")
```

```{r}
summary(lm(temp ~ date, data = sst_filter))
```

```{r}
# Random exploration 
#sst_ma <- sst_filter %>% mutate(ma = slider::slide_dbl(temp, mean,
                #.before = 2, .after = 2, .complete = TRUE))

# Plot the original time series
#ggplot() +
  #geom_line(data = sst_filter, aes(x = date, y = temp), color = "blue") +
  #geom_line(data = sst_ma, aes(x = date, y = ma), color = "red") +
  #labs(title = "Time Series with Moving Average")

```

```{r}
acf(sst_filter$temp, lag.max=12)
```

```{r}
#sst_ma <- sst_ma %>%
  #mutate(datetime = as_datetime(sst_filter$date))

#df_averaged <- sst_ma %>%
  #group_by(datetime) %>%
  #summarise(mean_value = mean(temp))


#decomp = as_tsibble(df_averaged)

#class(df_averaged$mean_values)
```

## Wrangling - sea surface temp data

-   changed date to datetime

-   got rid of na

-   arranged by index (date)

-   added index and temp to sst_filter df

-   selected for site1 (stearns warf) for SST1

-   added index to SST_filter 1

-   omitted NA for SST filter

```{r}
sst_filter <- sst_filter %>%
  mutate(index = as.Date(date))

sst_filter <- na.omit(sst_filter)

sst_filter <- sst_filter %>%
  arrange(index)

sst_filter <- sst_filter %>%
  mutate(index = as.Date(date),
         temp = as.numeric(temp))

sst_site1 <- sst_filter %>%
  filter(site == "SITE1")

sst_filter1 <- sst_site1 %>%
  mutate(index = as.Date(date))

sst_filter1 <- na.omit(sst_filter1)

sst_filter <- sst_filter1 %>%
  arrange(index)

sst_filter1 <- sst_filter1 %>%
  mutate(index = as.Date(date),
         temp = as.numeric(temp))


```

## Wrangling algae data

-   grouped by pseudo nitxschia seriata group

-   selected for stearns warf

-   

```{r}
names_algae <- algal_occurence %>%
  group_by(organismName) %>%
  na.omit()

pseudo_algae <- algal_occurence %>%
  filter(organismName == "Pseudo_nitzschia_seriata_group")

join <- left_join(pseudo_algae, algal_events, by = "id")

join_stearns <- join %>%
  filter(locationID == "HABs-StearnsWharf") %>%
  mutate(year = format(eventDate, "%Y"))

date_join_stearns <- left_join(join_stearns, updated_sst, by ="year")

join <- join[complete.cases(join$decimalLatitude, join$decimalLongitude), ]
join_sf <- st_as_sf(join, coords = c("decimalLatitude", "decimalLongitude"), crs = 4326)

ggplot(data = join_stearns, aes(x = eventDate, y = organismQuantity)) +
  geom_point() +
  
  


```

```{r}
#ggplot() +
  #geom_line(data = stearns_ma, aes(x = eventDate, y = organismQuantity), color = "blue") +
  #geom_line(data = stearns_ma, aes(x = eventDate, y = ma), color = "red") +
  #labs(title = "Time Series with Moving Average")

#class(stearns_ma$eventDate)

#decomp_stearn <- date_join_stearns %>%
  #mutate(y_m = yearmonth(date)) %>%
  #group_by(y_m) %>%
  #summarise(total_quant = sum(organismQuantity, na.rm = TRUE)) %>%
  #mutate(total_quant = )
  #as_tsibble(index = y_m)
  
```

```{r}
date_join_stearns <- date_join_stearns %>%
  mutate(y_m = yearmonth(date)) %>%
  filter(!is.na(date))

```

```{r}
decomp_stearn <- date_join_stearns %>%
  filter(!is.na(date)) %>%
  mutate(y_m = yearmonth(date)) %>%
  group_by(y_m) %>%
  summarise(total_quant = sum(organismQuantity, na.rm = TRUE)) %>%
  mutate(total_quant = na.approx(total_quant)) %>%
  as_tsibble(index = y_m)
```

```{r}
decomp_stearn %>% model(
    STL(total_quant)
  ) %>%
  components() %>%
  autoplot()+
  labs(title = "Stearns Warf, S")
```

```{r}
ggplot(data = date_join_stearns, aes(x = ))
```

```{r}
total_data <- left_join(date_join_stearns, decomp_stearn, by = "y_m")
```

```{r}
ggplot(data = total_data, aes(x = temp, y = total_quant)) +
  geom_line()
```

```{r}
summary(lm(organismQuantity ~ date, data = date_join_stearns))
```

```{r}
summary(lm(organismQuantity ~ y_m, data = date_join_stearns))
```

```{r}
summary(lm(total_quant ~ y_m, data = decomp_stearn))
```

```{r}
summary(lm(organismQuantity ~ y_m + temp, data = date_join_stearns))
```

```{r}
summary(lm(total_quant ~ temp, data = total_data))

```

```{r}

#mod <- lm(organismQuantity ~ temp + year, data = date_join_stearns)

date_join_stearns %>% 
  ggplot(aes(x = temp, y = organismQuantity, color = year)) +
  geom_point() +
  geom_line(data = augment(mod), aes(y = .fitted, color = year)) + 
  labs(x = "Engine Displacement, in litres",
       y = "Highway Miles per gallon") +
  scale_colour_discrete("year")
```

```{r}
ggplot(data = date_join_stearns, aes(x = temp)) +
  geom_histogram()
```

```{r}
date_join_stearns %>%
  group_by(y_m) %>%
  summarize(average_cells_per_liter = mean(organismQuantity, na.rm = TRUE))

```

```{r}
stearns_15 %>%
  group_by(y_m) %>%
  summarize(average_cells_per_liter = mean(organismQuantity, na.rm = TRUE))

```

```{r}
stearns_23 <- date_join_stearns %>%
  filter(year == "2022")

stearns_23 %>%
  group_by(y_m) %>%
  summarize(average_cells_per_liter = mean(organismQuantity, na.rm = TRUE))


```

```{r}

acf(decomp_stearn$total_quant, lag.max=52)

```

```{r}
res = feols(organismQuantity ~ temp + y_m, date_join_stearns)
# You can specify clustered standard-errors in summary:
summary(res, cluster = ~Species)res = feols(Sepal.Length ~ Sepal.Width + Petal.Length, iris)
# You can specify clustered standard-errors in summary:
summary(res, cluster = ~Species)
```

```{r}
date_join_stearns <- date_join_stearns %>%
  arrange(y_m) %>%
  mutate(lagged_temp = lag(temp, 1))

# Plot the lagged effect
ggplot(date_join_stearns, aes(x = lagged_temp, y = organismQuantity)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Lagged Temperature", y = "Number of Algae") +
  ggtitle("Lagged Effect of Temperature on Algae Growth")
```

```{r}
date_join_stearns <- date_join_stearns %>%
  mutate(year = format(y_m, "%Y"))
```

```{r}
ggplot(data = date_join_stearns, aes(x = y_m ))
```

```{r}
#California data wrangling 
usa <- map_data("usa") # we already did this, but we can do it again
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group)) + 
  coord_quickmap()

state <- map_data("state")

ca <- state %>%
  filter(region %in% c("california"))

ggplot(data = ca) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "black") + 
  coord_quickmap()

counties <- map_data("county")
ca_county <- counties %>%
  filter(region == "california")

ca_counties <- ggplot(data = ca_county) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "black") + 
  coord_quickmap()

ca_counties

```

```{r}

stearns_15 <- date_join_stearns %>%
  filter(year == "2015")


stearns_15 <- stearns_15 %>%
  mutate(month = format(date, "%m"))

#ggplot(data = stearns_15, aes(x = ))



```

```{r}
ggplot(data = stearns_15, aes(x = date, y = organismQuantity)) +
  geom_point()
```

```{ggplot(data = stearns_15, aes(x = y_m)) +}
  geom_line(aes(y = organismQuantity, color = "Line 1"), size = 1.5) +
  geom_line(aes(x = temp, y = organismQuantity, color = "Line 2"), size = 1.5) +
  labs(title = "Two Lines with Different X Variables") +
  xlab("Common X-axis Label") +
  ylab("Y-axis Label") +
  scale_color_manual(values = c("Line 1" = "blue", "Line 2" = "red"))


```
